---
title: "HW1 Question 1"
author: "Saksham Goel"
date: "February 3, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1 - Box Office Ticket Sales

### Introducing the Dataset

This section is dedicated to see and understand the data that was provided from the weekly reports about the box office ticket sales for plays in Broadway in New York. The data being observed is of the week of October 11- 17, 2004. The dataset contains data about the gross box office results for the current week October 11-17, 2004 and that of the previous week October 3-10, 2004.

The data table is as follows:
```{r q1_view_data}
ticketSales = read.csv("playbill.csv", header=T)
ticketSales[,]
```

### Visualizing the Dataset
The dataset can be visalized through a scatterplot in the figure below:
```{r q1_plot_data}
plot(x = ticketSales$LastWeek, y = ticketSales$CurrentWeek)
```

### Fitting a linear model
The linear trend in the scatterplot seems strong, which means that a linear regression model is appropriate. The linear model that we are trying to fit is of the form:

\begin{center}$\hat{Y} = \hat{\beta_0} + \hat{\beta_1}X$\end{center}

here,
X = Gross Box Office results for Previous Week
Y = Gross Box Office results for Current Week. 

Through this linear model we are trying to predict the value of actual gross box office reults of current week (Y) using the gross box office results of the previous week (X). The following snippet of the R-Code fits a linear model on the data, prints out the values of the intercept and slope and also provides a summary of the fitted model.
```{r q1_fit_model}
mod = lm( ticketSales$CurrentWeek ~ ticketSales$LastWeek )
mod
```

After fitting the model and observing the values of the parameters $\hat{\beta_0}, \hat{\beta_1}$ we find that:

\begin{center} $ \hat{\beta_0} = 6804.8860 $ \end{center}
\begin{center} $\hat{\beta_1} = 0.9821$ \end{center}


The summary of the model is as follows:
```{r q1_fit_model2}
summary(mod)
```
The summary provides us with a lot of useful values that we will use in the upcoming sections to calculate the Confidence Intervals, Prediction Intervals and perform Hypothesis Testing. Some of the values are as follows:

Sum of Residuals Square (s) = 18010\\
Degrees of freedom (n-2) = 16\\


Combining all the information from the previous sections the final fitted model looks as follows:

\begin{center}$Y = 6804.8860 + (0.9821 * X)$\end{center}

### Finding a 95% Confidence Interval for $\beta_1$

A $100(1-\alpha)\%$ confidence interval for $\hat{\beta_1}$ is given by the following formula:

\begin{center}
\begin{math}
\hat{\beta_{1}} -  t_{ \frac{\alpha}{2},n-2}\frac{s}{ \sqrt{ S_{XX} } } \leq \beta_1 \leq \hat{\beta_{1}} + t_{ \frac{\alpha}{2},n-2}\frac{s}{ \sqrt{ S_{XX} } }
\end{math}
\end{center}

The following snippet of code is used to find the 95% confidence interval for $\hat{\beta_1}$:

```{r q1_CI}
bet1_hat = 0.9821 #found using the linear model
x_col = ticketSales$LastWeek
y_col = ticketSales$CurrentWeek
x_bar = mean(x_col)
#x_bar = 622186.6
y_bar = mean(y_col)
#y_bar = 617842.8
sxx = sum((x_col-x_bar)^2)
#sxx = 1.557916 * 10^12
sxy = sum((x_col-x_bar)*(y_col-y_bar))
#sxy = 1.53 * 10^12
#sxy/sxx = 0.9821
s = 18010 #found through the summary of the linear model
t_mult = 2.120 # found using the T Table, equal to t0.25 for 16 degrees of freedom
beta1CIlower = bet1_hat - (t_mult*(s/sqrt(sxx)))
beta1CIupper = bet1_hat + (t_mult*(s/sqrt(sxx)))
```

The values we found are as follows:

\begin{math} \bar{x} = 622186.6 \end{math}\
\begin{math} \bar{y} = 617842.8 \end{math}\
\begin{math} S_{XX} = 1.557916 * (10 ^12) \end{math}\
\begin{math} S_{XY} = 1.53 * (10 ^12) \end{math}\
\begin{math} \hat{\beta_{1}} = 0.9820815 \end{math}\




The $95\%$ confidence interval can thus be given as follows: 
\begin{center}
\begin{math}
0.9515101 \leq \beta_{1} \leq 1.01269
\end{math}
\end{center}

Yes we can say that $1$ is a reasonable value for $\beta_1$ because $1$ lies in the $95\%$ Confidence Interval of $\beta_1$

### Hypothesis Testing for $\beta_0$
We need to run a Hypothesis Test for $\beta_0$ given by:
\begin{center}
\begin{math}
Null \big(H_{0}\big) \beta_{0} = 10000
Alternate \big(H_{1}\big) \beta_{0} \ne 10000
\end{math}
\end{center}

The hypothesis test for $\beta_0$ is given by the following formula:

\begin{center}
\begin{math}
\hat{T} = \frac{\hat{\beta_0} - \beta_0}{s\sqrt{\frac{1}{n} + \frac{\bar{x}^2}{S_{XX}}}}
\end{math}
\end{center}

To find the p-value we then use the t-statistic we got from the previous formula and find the probability as follows:
\begin{center}
\begin{math}
p-val = 2 \cdot P\big( t \geq \mid\hat{T}\mid\big)
\end{math}
\end{center}

The following snippet of code is used to perform the Hypothesis Test for $\beta_0$:

```{r q1_HT}
bet0_hat = 6804.8860 #this value was found using the linear model through r in the previous section
x_col = ticketSales$LastWeek
x_bar = mean(x_col)
#x_bar = 622186.6
sxx = sum((x_col-x_bar)^2)
#sxx = 1.557916 * 10^12
s = 18010 #found through the summary of the linear model
t_mult = 2.120 # found using the T Table, equal to t0.25 for 16 degrees of freedom
n = length(x_col)
beta0test = ((bet0_hat - 10000)/(s * sqrt((1/n) + ((x_bar^2)/sxx))))
#beta0test =  -0.3217422
p_val = 2*pt(-abs(beta0test), df = n-2)
#p_val = 0.7518132
```

After performing the Hypothesis test for $\beta_0$ we find that the $p-value = 0.7518132$ which is greater than $0.05$, hence we cannot reject the null hypothesis.

### Point Estimate for new Y

Through the previous sections we know that:
\begin{center}$Y = 6804.8860 + (0.9821 * X)$\end{center}

So to find the point estimate of Y using the fitted model we get:
```{r point_Estimate}
X = 400000 #X value for which we need to find the estimated value
Y = 6804.8860 + (0.9821 * X)
#Y = 399644.9
```

Through this above snippet of code we get the point estimate of $Y$ at $X = 400,000$ as $Y = 399644.9$.


### Prediction Interval for new Y
There is no use of a Linear Model if we cannot predict new values of the parameter Y through X. In this section we will construct a prediction interval of the Y value using the formulas provided in the book. The prediction interval of Y is given by the formula as follows:

\begin{center}
\begin{math}
\hat{y}_{n+1} \pm t_{\alpha/2,n-2}\sqrt{MSE} \sqrt{1+\dfrac{1}{n}+\dfrac{(x_{n+1}-\bar{x})^2}{\sum(x_i-\bar{x})^2}}
\end{math}
\end{center}

The following snippet of code helps us to find the upper and lower limits of the Prediction Interval for the new Y as follows:

```{r q1_PI}
#fitted model = 
# Y = 6804.8860 + 0.9821 * X
x_curr = 400000 #the value at which we need to find the prediction interval
y_hat = 6804.8860 + (0.9821 * x_curr)
#y_hat = 399644.9.
mse = mean(mod$residuals^2)  # mean((ticketSales$CurrentWeek - predict(mod))^2)
#mse = 288241878
x_col = ticketSales$LastWeek
x_bar = mean(x_col)
#x_bar = 622186.6
sxx = sum((x_col-x_bar)^2)
#sxx = 1.557916 * 10^12
s = 18010 #found through the summary of the linear model
t_mult = 2.120 # found using the T Table, equal to t0.25 for 16 degrees of freedom
n = length(x_col)
yPILower = y_hat - (t_mult * sqrt(mse) * sqrt(1 + (1/n) + (((x_curr - x_bar)^2)/sxx) ))
yPIUpper = y_hat + (t_mult * sqrt(mse) * sqrt(1 + (1/n) + (((x_curr - x_bar)^2)/sxx) ))
#yPILower = 362115
#yPIUpper = 437174.8
```

Through the above snippet of code the resulting $95%$ prediction interval is given as follows:

\begin{center}
\begin{math}
362115 \leq Y \leq 437174.8
\end{math}
\end{center}

Because the prediction interval does not contain the value $450,000$ in it, we cannot use that value as a prediction.

### Comment:
I think that is a statement that could be used frequently to describe the sales record for the broadway show, because the linear model also seems to follow a trend as seen. Other than that all the results from the previous computations seems to provide a good enough evidence to support this claim.