---
title: "HW1 Question 2"
author: "Saksham Goel"
date: "February 3, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 2 - Processing Time of Invoices

### Introducing the Dataset

This section is dedicated to see and understand the data that was provided from a large company about the number of invoices processed and the total processing time for those invoices. The data being observed is of a whole month. 
The data table is as follows:
```{r q1_view_data}
invoiceTime = read.table("invoices.txt", header = T)
invoiceTime[,]
```


### Visualizing the Dataset
The dataset can be visalized through a scatterplot in the figure below:
```{r q1_plot_data}
plot(x = invoiceTime$Invoices, y = invoiceTime$Time)
```

### Fitting a linear model
The linear trend in the scatterplot seems strong, which means that a linear regression model is appropriate. The linear model that we are trying to fit is of the form:

\begin{center}$\hat{Y} = \hat{\beta_0} + \hat{\beta_1}X$\end{center}

here,
X = Number of Invoices
Y = Total Time required to process those Invoices

Through this linear model we are trying to predict the value of total time required to process invoices (Y) using the number of invoices (X). The following snippet of the R-Code fits a linear model on the data, prints out the values of the intercept and slope and also provides a summary of the fitted model.
```{r q1_fit_model}
mod = lm(invoiceTime$Time ~ invoiceTime$Invoices )
mod
```

After fitting the model and observing the values of the parameters $\hat{\beta_0}, \hat{\beta_1}$ we find that:

\begin{center} $ \hat{\beta_0} = 0.64171 $ \end{center}
\begin{center} $\hat{\beta_1} = 0.01129 $ \end{center}


The summary of the model is as follows:
```{r q1_fit_model2}
summary(mod)
```
The summary provides us with a lot of useful values that we will use in the upcoming sections to calculate the Confidence Intervals, Prediction Intervals and perform Hypothesis Testing. Some of the values are as follows:

Sum of Residuals Square (s) = 0.3298\\
Degrees of freedom (n-2) = 28\\


Combining all the information from the previous sections the final fitted model looks as follows:

\begin{center}$Y = 0.6417099 + (0.0112916 * X)$\end{center}


### Finding a 95% Confidence Inrerval $\beta_0$
A $100(1-\alpha)\%$ confidence interval for $\hat{\beta_0}$ is given by the following formula:

\begin{center}
\begin{math}
\hat{\beta_{0}} -  t_{ \frac{\alpha}{2},n-2}\cdot s \cdot \sqrt{\frac{1}{n} + \frac{\bar{x}^2}{S_{XX}}} \leq \beta_0 \leq \hat{\beta_{0}} +  t_{ \frac{\alpha}{2},n-2}\cdot s \cdot \sqrt{\frac{1}{n} + \frac{\bar{x}^2}{S_{XX}}}
\end{math}
\end{center}

The following snippet of code is used to find the $95\%$ confidence interval for $\hat{\beta_0}$:

```{r q2_CI}
bet0_hat = 0.6417099 #using the linear model fit from prevous section
x_col = invoiceTime$Invoices
x_bar = mean(x_col)
#x_bar = 130.0333
sxx = sum((x_col-x_bar)^2)
#sxx = 162367
n = length(x_col)
#n = 30
s = 0.3298 #found through the summary of the linear model
t_mult = 2.048 #found using the T Table, equal to t0.25 for 28 degrees of freedom
beta0CIlower = bet0_hat - ( t_mult * s * sqrt( (1/n) + ((x_bar^2)/(sxx)) ) )
beta0CIupper = bet0_hat + ( t_mult * s * sqrt( (1/n) + ((x_bar^2)/(sxx)) ) )
#beta0CIlower = 0.3912792
#beta0CIupper = 0.8921406
```

The $95\%$ confidence interval can thus be given as follows: 
\begin{center}
\begin{math}
0.3912792 \leq \beta_{0} \leq 0.8921406
\end{math}
\end{center}


### Hypothesis Testing for $\beta_1$
We need to run a Hypothesis Test for $\beta_0$ given by:
\begin{center}
\begin{math}
Null \big(H_{0}\big) \beta_{1}= 0.01
Alternate \big(H_{1}\big) \beta_{1} \ne 0.01
\end{math}
\end{center}

The hypothesis test for $\beta_0$ is given by the following formula:

\begin{center}
\begin{math}
\hat{T} = \frac{\hat{\beta_1} - \beta_1}{\frac{s}{\sqrt{S_XX}}}
\end{math}
\end{center}

To find the p-value we then use the t-statistic we got from the previous formula and find the probability as follows:
\begin{center}
\begin{math}
p-val = 2 \cdot P\big( t \geq \mid\hat{T}\mid\big)
\end{math}
\end{center}

The following snippet of code is used to perform the Hypothesis Test for $\beta_1$:

```{r q1_HT}
bet1_hat = 0.0112916 #found using the linear model in the previous sections
x_col = invoiceTime$Invoices
y_col = invoiceTime$Time
x_bar = mean(x_col)
#x_bar = 130.0333
y_bar = mean(y_col)
#y_bar = 2.11
sxx = sum((x_col-x_bar)^2)
#sxx = 162367
sxy = sum((x_col-x_bar)*(y_col-y_bar))
#sxy = 1833.39
#sxy/sxx = 0.0112916
n = length(x_col)
s = 0.3298 #found through the summary of the linear model
t_mult = 2.048 #found using the T Table, equal to t0.25 for 28 degrees of freedom
beta1test = ( ( bet1_hat - 0.01 )/( s/sqrt( sxx ) ) )
#beta1test = 1.57807
p_val = 2*pt(-abs(beta1test), df = n-2)
#p_val = 0.1257819
```

After performing the Hypothesis test for $\beta_1$ we find that the $p-value = 0.1257819$ which is greater than $0.05$, hence we cannot reject the null hypothesis.

### Point Estimate for new Y

Through the previous sections we know that:
\begin{center}$Y = 0.6417099 + (0.0112916 * X)$\end{center}

So to find the point estimate of Y using the fitted model we get:
```{r point_Estimate}
X = 130 #X value for which we need to find the estimated value
Y = 0.6417099 + 0.0112916 * X
#Y = 2.109618
```

Through this above snippet of code we get the point estimate of $Y$ at $X = 130$ as $Y = 2.109618$ which is roughly equal to $127 mins$.


### Prediction Interval for new Y
There is no use of a Linear Model if we cannot predict new values of the parameter Y through X. In this section we will construct a prediction interval of the Y value using the formulas provided in the book. The prediction interval of Y is given by the formula as follows:

\begin{center}
\begin{math}
\hat{y}_{n+1} \pm t_{\alpha/2,n-2}\sqrt{MSE} \sqrt{1+\dfrac{1}{n}+\dfrac{(x_{n+1}-\bar{x})^2}{\sum(x_i-\bar{x})^2}}
\end{math}
\end{center}

The following snippet of code helps us to find the upper and lower limits of the Prediction Interval for the new Y as follows:

```{r q1_PI}
#fitted model = 
# Y = 0.6417099 + 0.0112916 * X
x_curr = 130 #the value at which we need to find the prediction interval
y_hat = 0.6417099 + (0.0112916 * x_curr)
#y_hat = 2.109618
mse = mean(mod$residuals^2) # mean((invoiceTime$Times - predict(mod))^2)
#mse = 0.1015004
x_col = invoiceTime$Invoices
x_bar = mean(x_col)
#x_bar = 130.0333
sxx = sum((x_col-x_bar)^2)
#sxx = 162367
n = length(x_col)
s = 0.3298 #found through the summary of the linear model
t_mult = 2.048 #found using the T Table, equal to t0.25 for 28 degrees of freedom
yPILower = y_hat - (t_mult * sqrt(mse) * sqrt(1 + (1/n) + (((x_curr - x_bar)^2)/sxx) ))
yPIUpper = y_hat + (t_mult * sqrt(mse) * sqrt(1 + (1/n) + (((x_curr - x_bar)^2)/sxx) ))
#yPILower = 1.446357
#yPIUpper = 2.772878
```

Through the above snippet of code the resulting $95%$ prediction interval is given as follows:

\begin{center}
\begin{math}
1.446357 \leq Y \leq 2.772878
\end{math}
\end{center}
