---
title: "Lab 3 Worksheet"
author: "Saksham Goel"
date: "January 30, 2018"
output: pdf_document
---

# The Timing of Production Runs.  
We have talked about the study of production runs during the lecture. Now it is time for you to explore the data firsthand. 

## Download data
Download the data **production.txt** from the textbook website <www.stat.tamu.edu/~shether/book>

## Read data into R
Read the data into R by using function "read.table( )" and save the data in a variable called "prod". Prod will be a type of R object called data frame. 

You can provide a complete address for the file or set the working directory to the folder where you have saved the data:

prod = read.table('/myComputer/myFolder1/myFolder2/production.txt', header = TRUE)

Or

setwd('/myComputer/myFolder1/myFolder2') 

prod = read.table('production.txt', header = TRUE)

```{r download_data}
# read data into R. 
prod = read.table("production.txt", header= TRUE)
# Use function View(prod) or head(prod) to see if the data has been imported successfully. 
#View(prod)
head(prod)
# Advanced: can you try importing data using "read.csv( )" instead of "read.table( )":
csv_type = read.csv("production.txt", header = TRUE, sep = "\t")
#View(csv_type)
head(csv_type)
```

## Explore data 
**summary( )** is a very important function. See what happens when you apply it to a dataframe. 

```{r explore_data1}
# explore data using summary()
summary(prod)

```
Now let us draw some exploratory plots. 

For each variable in the data frame, we can draw a histogram. A histogram can give us an idea of the distribution of a variable. 

```{r explore_data2}
# explore the data with histograms
# Hint: use hist() for histograms
hist(prod$RunTime)
hist(prod$RunSize)
```

Next we draw a scatterplot with one variable (RunSize) on the x axis and another variable (RunTime) on the y axis. 

```{r explore_data3}
# explore data with scatterplot
# Hint: use plot( )
plot(x = prod$RunSize, y = prod$RunTime)
```

## Fit a regression model
The linear trend in the scatterplot seems strong, which means that a linear regression model is appropriate. 

```{r fit_model1}
# fit a linear regression model
# Hint: use lm( ). Don't forget to save the model to a variable called "mod"
mod = lm( prod$RunTime ~ prod$RunSize )
```

We have applied the function summary( ) to a data frame. This function can also be applied to a model. 

```{r fit_model2}
# apply summary( ) to your model
summary(mod)
```

Please complete the following formula of the model:

Fitted RunTime = 149.74770 + 0.25924 * RunSize

Or we can use mathematical symbols:

Let X = RunSize, Y = RunTime. $\hat{Y} = \hat{\beta_0} + \hat{\beta_1}X$

## Compare the observed RunTime and the fitted RunTime

```{r compare}
# Print out the observed RunTime
# prod$RunTime 
# Print out the observed RunTime
# Hint: Use mod$fitted.values OR use predict(mod)
prod$RunTime
predict(mod)

```

The observed RunTime is different from the fitted RunTime. 

## visualize the observed RunTime and the fitted RunTime

```{r visual}
# Hint: first draw a scatterplot of the data
plot(x = prod$RunSize, y = prod$RunTime)
# Next add the fitted values
points(predict(mod) ~ prod$RunSize, col = 'red', pch = 4)
```

## Calculating the estimated coefficients using formulas
We learnt that the following formulas during the lecture:

$$\hat{\beta_1} = \frac{\sum_{i=1}^n (x_i-\bar{x})(y_i-\bar{y})}{\sum_{i=1}^n (x_i-\bar{x})^2} $$

$$\hat{\beta_0} = \bar{y} - \hat{\beta_1}\bar{x}$$

Let's use these two formulas to estimate the coefficients. 
```{r calculate}
 x = prod$RunSize
 y = prod$RunTime
 xbar = mean(x)
 ybar = mean(y)
 
 beta1_hat = sum((x-xbar)*(y-ybar)) / sum((x-xbar)^2)
 beta1_hat
 beta0_hat = ybar - beta1_hat*xbar
 beta0_hat
```
Compare beta1_hat and beta2_hat with the estimated coefficients obtained from the linear regression model. They should be identical! 



\\
*Congratulations! You’ve fitted a simple linear regression model. Please generate the pdf and wait for further instruction. *