---
title: "R Exercise2"
author: "Georgia Huang"
date: "1/23/2018"
output: pdf_document
---

## A Simple Data Analysis
We will study the hurricanes dataset, consisting of 287 hurricanes that had made landfall in the US between 1851 and 2013. 


### Step 1: import data
```{r step1}
setwd('/Users/zhuozhihuang/Desktop')  # set the director/folder you want to work in. 
hur = read.csv('hurricanes.csv')
hur1 = hur[, 1:5] # we only want the first 5 columns
```

### Step 2: explore data 
Information about the variables:
* Year, Month: The year and month in which the hurricane made landfall.
* HighestSS: The highest Saffir–Simpson hurricane wind scale that the hurricane produced in the US.
* Pressure: The central pressure of the hurricane at landfall, measured in hPa.
* MaxWind: The highest sustained surface wind occurred along the US coast
* Name: The name of the hurricane, if available.

#### summary statistics
```{r step2.1}
head(hur1)  # view the first few lines of data
names(hur1) # the name of the variables (column headers)
dim(hur1) # how many rows, how many columns
summary(hur1)
```
#### exploratory plots
```{r step2.2}
hist(hur1$Pressure)
hist(hur1$MaxWind)
```
Histograms can give us some idea about the underlying probability distributions of the variables. 

### Step 3: Evaluate the relationship between MaxWind and Pressure
```{r step 3.1}
plot(Pressure~MaxWind, data = hur1)
```
It seems that as MaxWind increases, pressure tends to decrease.Let us add a straight line to fit the relationship. 


### Step 4: Output/save data
```{r step4}
write.csv(hur1, 'hurricanes_partial.csv')
```

## Simulation /Random sampling
R can create a random sample. Let us use get a random sample from Normal(2, 4). 

```{r step5}
a = rnorm(n = 100, mean = 2, sd = 2)
hist(a)
```
It looks somewhat symmetric with center at 2, but not exactly normal. This is because our sample size is not large enough. 

```{r step6}
a2 = rnorm(n = 1000, mean = 2, sd = 2)
hist(a2)
```
It is getting close to the shape of a normal distribution. Let's increase the sample size again. 

```{r step7}
a2 = rnorm(n = 10000, mean = 2, sd = 2)
hist(a2)
```
It lookes like a normal distribution centered at 2! If you increase the sample size, the approximation will get better and better. 